### 2022-02-13 prior_versions/revision_journal/setup.tex
update

First, we assume the row vectors of $X^{(i)}$ are i.i.d. centered random vectors with  population covariance matrix $\Sigma^{(i)} \in \real^{p \times p}$. Equation \eqref{eq_rvm} implies that
\be\label{XofZ}
    X^{(i)} = Z^{(i)} (\Sigma^{(i)})^{1/2} \in \real^{n_i\times p},
\ee
recalling that $\Sigma^{(i)}$ is a deterministic positive definite symmetric matrices, $Z^{(i)}=(z^{(i)}_{j,k})$ is a $n_i\times p$ random matrix with real-valued independent entries having zero mean and unit variance.
%\begin{equation}\label{assm1}
%\mathbb{E} z^{(1)}_{ij}=\mathbb{E} z^{(2)}_{ij} =0, \ \quad \ \mathbb{E} \vert z^{(1)}_{ij} \vert^2=\mathbb{E} \vert z^{(2)}_{ij} \vert^2 =1,
%\end{equation}
Let $\tau>0$ be a small constant.
We assume that for some constant $\varphi>4$, the $\varphi$-th moment of each entry $z^{(i)}_{j, k}$ is bounded from above by $1/\tau$:
\be \label{conditionA2}
\ex{\vert z^{(i)}_{j, k} \vert^\varphi}  \le \frac 1{\tau}. %,\quad \mathbb{E} \vert  z^{(2)}_{ij} \vert^\varphi   \le \tau^{-1}
\ee
%$Denote by $\Sig^{(i)} = O_i\Lambda_i O_i^\top$ its eigendecomposition, where $\Lambda_1=\text{diag}(\si_1^{(i)}, \ldots, \si^{(i)}_n)$.
%\be\label{eigen}
%\Sig^{(i)}, \quad  ,  \quad i=1,2,
%\ee
Additionally, We assume that the eigenvalues of $\Sigma^{(i)}$ are bounded between $\tau$ and $1 / \tau$.
%\begin{equation}\label{assm3}
%    \tau \le  \si^{(i)}_p \le\cdots\le \si^{(i)}_2 \le \si^{(i)}_1 \le \tau^{-1} \quad i=1,2.
%\end{equation}

@@

\paragraph{Information transfer.} To be precise, we say that task one provides a \textit{positive transfer} to task two, if the excess risk $L(\hat{\beta}_2^{\MTL})$ is smaller than $L(\hat{\beta}_2^{\STL})$.
Otherwise, we say that task one provides a \textit{negative transfer} to task two.

@@

We now illustrate that the bias-variance decomposition of Proposition \ref{lem_HPS_loss} characterizes whether or not the information transfer is positive.
First, we compare the bias. Since $n_1 \ge (1 + \tau) p$ by Assumption \ref{assm_big1}, the bias of $\hat{\beta}_2^{\STL}$ is zero.
Thus, the bias of $\hat{\beta}_2^{\MTL}$ is always higher than that of $\hat{\beta}_2^{\STL}$.
Second, we compare the variance. By a similar argument as in Proposition \ref{lem_HPS_loss}, the variance of $\hat{\beta}_2^{\STL}$ is equal to $\sigma^2 \bigtr{\Sigma^{(2)} ((X^{(2)})^{\top} X^{(2)})^{-1}}$.
Then, a straightforward algebraic calculation (e.g. using the Woodbury matrix identity) shows that the variance of $\hat{\beta}_2^{\MTL}$ is always higher than that of $\hat{\beta}_2^{\STL}$:
\begin{align*}
    \bigtr{\Sigma^{(2)} \big(\hat{\Sigma}(a)\big)^{-1}} \le \bigtr{\Sigma^{(2)} \big((X^{(2)})^{\top} X^{(2)}\big)^{-1}}.
\end{align*}
Combining both comparisons, we conclude that HPS affects the excess risk in two ways: the bias always increases while the variance always decreases.
Thus, whether or not task one provides a positive transfer to task two depends on which effect is more significant.

>>>>

First, recall that the row vectors of $X^{(i)}$ are i.i.d. centered random vectors with  population covariance matrix $\Sigma^{(i)} \in \real^{p \times p}$:
\be\label{XofZ}
    X^{(i)} = Z^{(i)} (\Sigma^{(i)})^{1/2} \in \real^{n_i\times p}.
\ee
%recalling that $\Sigma^{(i)}$ is a deterministic positive definite symmetric matrices,
where $Z^{(i)}=(z^{(i)}_{j,k})$ is a $n_i\times p$ random matrix with real-valued independent entries having zero mean and unit variance.
%\begin{equation}\label{assm1}
%\mathbb{E} z^{(1)}_{ij}=\mathbb{E} z^{(2)}_{ij} =0, \ \quad \ \mathbb{E} \vert z^{(1)}_{ij} \vert^2=\mathbb{E} \vert z^{(2)}_{ij} \vert^2 =1,
%\end{equation}
Let $\tau>0$ be a small constant.
We assume that for some constant $\varphi>4$, the $\varphi$-th moment of each entry $z^{(i)}_{j, k}$ is bounded from above by $1/\tau$:
\be \label{conditionA2}
\ex{\vert z^{(i)}_{j, k} \vert^\varphi}  \le \frac 1{\tau}. %,\quad \mathbb{E} \vert  z^{(2)}_{ij} \vert^\varphi   \le \tau^{-1}
\ee
%$Denote by $\Sig^{(i)} = O_i\Lambda_i O_i^\top$ its eigendecomposition, where $\Lambda_1=\text{diag}(\si_1^{(i)}, \ldots, \si^{(i)}_n)$.
%\be\label{eigen}
%\Sig^{(i)}, \quad  ,  \quad i=1,2,
%\ee
Additionally, We assume that the eigenvalues of $\Sigma^{(i)}$ are bounded between $\tau$ and $1 / \tau$.
%\begin{equation}\label{assm3}
%    \tau \le  \si^{(i)}_p \le\cdots\le \si^{(i)}_2 \le \si^{(i)}_1 \le \tau^{-1} \quad i=1,2.
%\end{equation}

@@

%\paragraph{Information transfer.} To be precise, we say that task one provides a \textit{positive transfer} to task two, if the excess risk $L(\hat{\beta}_2^{\MTL})$ is smaller than $L(\hat{\beta}_2^{\STL})$.
%Otherwise, we say that task one provides a \textit{negative transfer} to task two.

@@

We now illustrate that the bias-variance decomposition of Proposition \ref{lem_HPS_loss} characterizes whether or not the information transfer is positive.
First, we compare the bias. Since $n_1 \ge (1 + \tau) p$ by Assumption \ref{assm_big1}, the bias of $\hat{\beta}_2^{\STL}$ is zero.
Thus, the bias of $\hat{\beta}_2^{\MTL}$ is always higher than that of $\hat{\beta}_2^{\STL}$.
Second, we compare the variance. By a similar argument as in Proposition \ref{lem_HPS_loss}, the variance of $\hat{\beta}_2^{\STL}$ is equal to $\sigma^2 \bigtr{\Sigma^{(2)} ((X^{(2)})^{\top} X^{(2)})^{-1}}$.
Then, a straightforward algebraic calculation (e.g. using the Woodbury matrix identity) shows that the variance of $\hat{\beta}_2^{\MTL}$ is always higher than that of $\hat{\beta}_2^{\STL}$:
\begin{align*}
    \bigtr{\Sigma^{(2)} \big(\hat{\Sigma}(a)\big)^{-1}} \le \bigtr{\Sigma^{(2)} \big((X^{(2)})^{\top} X^{(2)}\big)^{-1}}.
\end{align*}
Combining both comparisons, we conclude that the transfer effect of HPS is mixed: the bias always increases while the variance always decreases.
Thus, whether or not the resulting information transfer is \textit{positive} depends on which effect is more significant.

### 2022-02-13 prior_versions/writeup/extension.tex
update

\section{Extensions to the Sparse Setting}
One important case is the sparse case, where we assume that the $\beta_i$'s are sparse in the sense
that each $\beta_i$ is in a subspace of dimension $k_i\ll p$. In other words,
$$\beta_i = \sum_{j=1}^{k_i} a^{(i)}_j v^{(i)}_j,$$
where $\{v^{(i)}_j\}$ is a set of $k_i$ orthonormal vectors. For $X_i= Z_i \Sigma_i^{1/2}$, we have
$$ X_i \beta_i = Z_i \Sigma_i^{1/2}\beta_i=  Z_i \Sigma_i^{1/2}P_i\beta_i,\quad P_i:= \sum_{j=1}^{k_i}v_j^{(i)}(v_j^{(i)})^\top.$$
Then we have
$$\frac{1}{n_i}\mathbb E\left[P_i \Sigma_i^{1/2} Z_i^T Z_i \Sigma_i^{1/2}P_i\right]=P_i \Sigma_i P_i.$$
Hence our analysis will be valid with projected covariance matrices $P_i \Sigma_i P_i$. Moreover, due to the sparsity assumption, it is natural to have restricted isometry property:
$$\frac{1}{n_i} P_i \Sigma_i^{1/2} Z_i^T Z_i \Sigma_i^{1/2}P_i \approx P_i \Sigma_i P_i.$$
Finally, for our analysis it is not necessary to assume that $\beta_T$ is sparse for the target.

>>>>

\section{Extensions to the Sparse Setting}

\todo{Describe a straightforward extension of our results here through isometry.}
One important case is the sparse case, where we assume that the $\beta_i$'s are sparse in the sense
that each $\beta_i$ is in a subspace of dimension $k_i\ll p$. In other words,
$$\beta_i = \sum_{j=1}^{k_i} a^{(i)}_j v^{(i)}_j,$$
where $\{v^{(i)}_j\}$ is a set of $k_i$ orthonormal vectors. For $X_i= Z_i \Sigma_i^{1/2}$, we have
$$ X_i \beta_i = Z_i \Sigma_i^{1/2}\beta_i=  Z_i \Sigma_i^{1/2}P_i\beta_i,\quad P_i:= \sum_{j=1}^{k_i}v_j^{(i)}(v_j^{(i)})^\top.$$
Then we have
$$\frac{1}{n_i}\mathbb E\left[P_i \Sigma_i^{1/2} Z_i^T Z_i \Sigma_i^{1/2}P_i\right]=P_i \Sigma_i P_i.$$
Hence our analysis will be valid with projected covariance matrices $P_i \Sigma_i P_i$. Moreover, due to the sparsity assumption, it is natural to have restricted isometry property:
$$\frac{1}{n_i} P_i \Sigma_i^{1/2} Z_i^T Z_i \Sigma_i^{1/2}P_i \approx P_i \Sigma_i P_i.$$
Finally, for our analysis it is not necessary to assume that $\beta_T$ is sparse for the target.
